---
title: "Day 3"
author: "BHV"
date: "2023-08-23"
output: html_document
---

## Exercises pt. 1
###1.Which carrier has the worst average delays? Challenge: can you disentangle the effects of bad airports vs. bad carriers? Why/why not?
```{r}
require(tidyverse)
require(nycflights13)

flights %>% group_by(carrier) %>% 
  summarize(avg_del_carrier = mean(arr_delay, na.rm = TRUE)) %>%
  arrange(desc(avg_del_carrier)) %>% 
  left_join(airlines) %>%
  select(avg_del_carrier,name,carrier)


# # second part
# 
# df1 <- flights %>% group_by(origin,dest,carrier) %>% 
#   summarize(avg_arr_delay_pr_carrier_pr_route = mean(arr_delay, na.rm = TRUE)) %>% arrange(carrier)
# 
# df2 <- flights %>% group_by(origin,dest)
# 
# %>% 
#   ungroup() %>% 
#   group_by(carrier) %>% 
#   summarize(avg_arr_delay_pr_carrier = sum(avg_arr_delay_pr_carrier_pr_route)) %>% arrange(desc(avg_arr_delay_pr_carrier))


```

###2.Find the flights that are most delayed upon departure from each destination.
```{r}
flights %>% group_by(origin) %>% slice_max(order_by = dep_delay) %>% select(origin, dep_delay, arr_delay)


```


###3.How do delays vary over the course of the day. Illustrate your answer with a plot.

```{r}
#Doesnt work
# flights %>% ggplot(aes(sched_dep_time,mean(!is.na(dep_delay)))) + geom_col(position = "dodge")

flights %>% group_by(cut(sched_dep_time,c(0,
                                          100,
                                          200,
                                          300, 
                                          400,
                                          500,
                                          600,
                                          700,
                                          800,
                                          900,
                                          1000,
                                          1100,
                                          1200,
                                          1300,
                                          1400,
                                          1500,
                                          1600,
                                          1700,
                                          1800,
                                          1900,
                                          2000,
                                          2100,
                                          2200,
                                          2300))
                     ) %>% summarise(avg_delay = mean(dep_delay, na.rm=TRUE,),
                                     number_of_flights = n()
                     ) %>% 
  ggplot(.,aes(.[[1]],avg_delay, fill = avg_delay)) +
  geom_col() +
  theme(axis.text.x = element_text(angle = 45))


flights %>% group_by(hour) %>% summarize(avg_delay = mean(dep_delay, na.rm = TRUE)) %>% 
  ggplot(aes(hour,avg_delay)) + 
  geom_col() +
  labs(title = "Avg. delay in minutes at each hour during the day",
       y = "Average delay in minutes") +
  theme_bw(base_size = 14)


flights %>% group_by(hour) %>% summarize(n_flights = n())
flights %>% group_by(hour) %>% summarize(avg_delay = mean(dep_delay, na.rm = TRUE),
                                         n_flights = n()) %>% 
  ggplot(aes(hour,avg_delay/n_flights)) + 
  geom_col() +
  labs(title = "Avg. delay in minutes per flight at each hour during the day",
       y = "Average delay per flight in minutes") + 
  theme_bw(base_size = 14)


```


###4.What happens if you supply a negative n to slice_min() and friends?


```{r}

#removes the last n observations

```
## Exercises pt. 2
### 1. montecarlo code
```{r}

#original

# Using this N
N <- 2e5

#fixing rng

set.seed(1)


monte_carlo <- function(N) {
  
  hits <- 0
  
  for (i in seq_len(N)) {
    x <- runif(1)
    y <- runif(1)
    if (y < x^2) {
      hits <- hits + 1
    }
  }
  
  hits / N
}


system.time(res <- monte_carlo(N))
res


# x and y outside
monte_carlo2 <- function(N) {
  
hits <- 0
    x = runif(N)
    y = runif(N)
    
  
  for (i in seq_len(N)) {
    if (y[[i]] < x[[i]]^2) {
      hits <- hits+1
    }
  }
  
  sum(unlist(hits)) / N
}
system.time(res <- monte_carlo2(N))
res

# take 3
monte_carlo3 <- function(N) {
  
output <- tibble(
    ex = runif(N),
    why = runif(N))
    
output <- output %>% mutate(below = ifelse(why < ex^2, 1L, 0L))
  
  
  sum(output[["below"]])/N
}
system.time(res <- monte_carlo3(N))
res

# Florian's solution

monte_carlo4 <- function(N) {
    mean(runif(N) < runif(N)^2)
}
system.time(monte_carlo4(N))
res

```
### ex 5.7
```{r}


# Start
system.time({
  set.seed(1)
  N <- 1e5
  x <- 0
  count <- 0
  for (i in seq_len(N)) {
    y <- rnorm(1)
    x <- x + y
    if (x < 0) count <- count + 1
  }
  p <- count / N
})

p


# v 2
rm(p,  x, count)

system.time({
  set.seed(1)
  N <- 1e5
  p <- tibble(y = rnorm(N)) %>% 
    mutate(x = cumsum(y)) %>% 
  filter(x < 0 )
  
  p <- sum(count(p["x"]))/N
})

p
rm(p,N)
# Florian's solution

system.time({
    set.seed(1)
    N <- 1e5
    y <- rnorm(N)
    x <- cumsum(y)
    p <- mean(x < 0)
})
p
```
#### pt 2
```{r}
mat <- as.matrix(mtcars)
ind <- seq_len(nrow(mat))
mat_big <- mat[rep(ind, 1000), ]  ## 1000 times bigger dataset
last_row <- mat_big[nrow(mat_big), ]

```

```{r}

#Original
system.time({
  for (j in 1:ncol(mat_big)) {
    for (i in 1:nrow(mat_big)) {
      mat_big[i, j] <- 10 * mat_big[i, j] * last_row[j]
    }
  }
})


# Florian's solution 

system.time({
    mat_big2 <- sweep(mat_big, 2, 10 * last_row, "*")
})
```
#### pt. 3
```{r}

#Start
M <- 50
set.seed(1)
step1 <- runif(M)
set.seed(1)
A <- rnorm(M)
N <- 1e4

tau <- matrix(0, N + 1, M)
tau[1, ] <- A

for (i in 2:nrow(tau)) {
for (j in 1:M) {
  
    tau[i, j] <- tau[i - 1, j] + step1[j] * 1.0025^(i - 2)
  }
} 



```

